Automated topic naming to support analysis of software maintenance activities

### 2010-08-20 icse/icse2011.tex
improved future work

% our field
%  Within the field of mining software repositories, a
To understand the latent subject matter of software development artifacts, researchers have proposed topic modeling and concept location.
%These techniques rely on the automated application of machine learning algorithms to corpora.
These techniques use unsupervised machine learning algorithms find topics or concepts.
% the problem
These topics are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
%Results are usually abstract, as the topics that are automatically extracted 
% the solution 
In order to have a meaningful discussion about the topics of software development, we must be able to interpret and devise labels for these topics. 
%However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
% the labelling research
Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
% a sub problem
% While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
% what we do

To solve this problem, this paper introduces \emph{labelled topic extraction}, 
in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
% data mining generalizable
Our approach is 
%To ensure we have a 
generalizable and project-agnostic, as we label topics using non-functional requirements,
% these qualities 
as these apply across many software systems.% and their development artifacts.
% We used real world data
We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
% We used LDA, CVS, Bitkeeper
We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their source control systems (CVS and BitKeeper). 
% We show NFRs exist
%Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, 
%and we identified topic trends over time. 	
% we can label stuff
Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, providing insight into their evolving software development activities.
\end{abstract}

@@

% why are topics useful
Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
Topics, too, are a valuable and automatically extractable abstraction of software artifacts.
Topics abstracts these artifacts by their subject matter, rather than their concrete representation. 
Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
The topic of development is relevant because it identifies the distinct issues facing development at a particular time.

@@

This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}). We show that these labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}). We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). The paper concludes with a discussion of limitations (Section \ref{sec:limit}) and related work (Section \ref{sec:related}).

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}). 
There is some debate about the significance and importance of the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research. 
The terms extracted from ISO9126 may not capture all words associated with the labels.  
For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
We therefore took the words from the taxonomy and expanded them.

@@

\noindent \textbf{Creating a validation corpus} --- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). 
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
We compared against this data-set, but we also used it for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

\subsection{Supervised labelling}
\label{sec:suplabelling}
Supervised labelling requires expert analysis of the correct word to assign to a topic. %In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.

@@

\begin{figure*}
\centering
\subfloat[MySQL]{
\includegraphics[width=0.4\textwidth]{figures/multi-label-results-mysql}
\label{fig:subfig3}
}
\subfloat[MaxDB]{
\includegraphics[width=0.4\textwidth]{figures/multi-label-results-maxdb}
\label{fig:subfig4}
}
\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.
%\todo[inline]{better discussion}% and graphical legend would be nice}
}
\label{fig:mulan}
\end{figure*}

@@

\subsection{Summary of techniques}
While unsupervised techniques such as LDA is appealing in its lack of human intervention, and thus lower effort, 
supervised learners have the advantage of domain knowledge which typically means improved results. 
Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.

@@

We observed that MySQL had more topic subsets than MaxDB. 
MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
MaxDB differed from MySQL since MaxDB was being actively developed, whereas halfway through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1. 
In other words, MySQL 3.23 was being maintained rather than actively developed, whereas MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
%\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}

@@

\textbf{Do label frequencies change over time?} -- Yes, MySQL's label
frequencies decreased as it got older. \emph{Usability} and \emph{reliability}
labels became more and more infrequent as it matured. \emph{Maintainability} topics
became more prevalent as MaxDB matured.

@@

%XXX can just remove this
\subsection{Effectiveness}

%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels, often the learners would mislabel a topic deserving of an infrequent label with a more common label.
The word-lists for \emph{correctness}, tended to be too lengthy, non-specific and broad, especially if WordNet words were used as it is a loosely defined concept in common parlance.

@@

\emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages. 
Conversely construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
We also chose our taxonomy and the data to study. Also, we did not show these results to the stakeholders.

\emph{Internal validity} issues are include inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or theory.

@@

% in Case Study Research: Design Case Studies
\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
Unfortunately only two annotators were used, their annotating could be biased and we did not analyze for inter-rater reliability.

@@

Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
They do not use external domain-specific information as we do.
In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, it does not leverage machine learning strategies.

@@

% para 1 we did this, we were successful
In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling latent software development topics using a non-functional requirements (NFRs) taxonomy. %topics and artifacts from project histories.
Our contributions include:
\begin{itemize}
\item We provide an unsupervised method of topic labelling (word-lists);
%\item Our unsupervised method can be applied to projects without training;
\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
\item We provide a method of cross-project analysis via topic labelling;
\item We demonstrate a method of visualizing these topics across time.
\end{itemize}

@@

% para 2 we did this to validate and show we did it
% Experiment description -- origin of wordbags
We validated our topic labelling techniques using multiple experiments.
Using unsupervised word-list analysis we set up two NFR topic labelling experiments: 
our \textsf{exp2} experiment used small but accurate word-lists, while \textsf{exp3} used large general word-lists derived from WordNet.
The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
%Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
% we could can improve auto labelling with ML
Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
Both kinds of learners performed well with area under the ROC curve values between $.6$ and $.8$.

% para 3 look while we did it we found this
% total topics - maybe move above
We showed that NFRs are often trending topics, that NFRs are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
% We found NFRs
Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project. 
% Try to provide an example of NFRs but little context
In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.

@@

\subsection{Future Work}

 % para 5 future work
 % Future work
There are several avenues of further investigation.  
 % Talk to people
We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
It is difficult to map abstract qualities to particular messages. % what?
 % This is disjointed and needs more context
On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance? %In this study we take the position that any type of maintenance has to do with software quality.

>>>>

% our field
%  Within the field of mining software repositories, a
To understand the latent subject matter of software development artifacts, researchers have proposed topic modeling and concept location.
%These techniques rely on the automated application of machine learning algorithms to corpora.
These techniques use unsupervised machine learning algorithms find topics or concepts.
% the problem
These topics are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
%Results are usually abstract, as the topics that are automatically extracted 
% the solution 
In order to have a meaningful discussion about the topics of software development, we must be able to attach meaningful labels to these topics. 
%However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
% the labelling research
Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
% a sub problem
% While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
% what we do

To solve this problem, this paper introduces \emph{labelled topic extraction}, 
in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
% data mining generalizable
Our approach is 
%To ensure we have a 
generalizable and project-agnostic, as we label topics using non-functional requirements,
% these qualities 
which apply across software systems.% and their development artifacts.
% We used real world data
We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
% We used LDA, CVS, Bitkeeper
We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their source control systems (CVS and BitKeeper). 
% We show NFRs exist
%Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, 
%and we identified topic trends over time. 	
% we can label stuff
Our labelled topic extraction technique produces appropriate, context-sensitive labels relevant to these projects, providing fresh insight into their evolving software development activities.
\end{abstract}

@@

% why are topics useful
Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
The topic of development is relevant because it identifies the distinct issues facing development at a particular time.

@@

This paper first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples. We begin by describing how we generated our data (Section \ref{sec:wordlist}). We show that labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}). We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). The paper concludes with a discussion of limitations (Section \ref{sec:limit}) and related work (Section \ref{sec:related}).

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}). 
There is some debate about the significance and importance of the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
We use these qualities later on as classes in supervised labelling.
The terms extracted from ISO9126 may not capture all words associated with the labels.  
For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
We therefore took the words from the taxonomy and expanded them.

@@

\noindent \textbf{Creating a validation corpus} --- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). 
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

\subsection{Supervised labelling}
\label{sec:suplabelling}
Supervised labelling requires expert analysis of the correct class/label to assign to a topic. In our approach, we use the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.

@@

\begin{figure*}[]
\centering
\subfloat[MySQL]{
\includegraphics[width=0.3\textwidth]{figures/multi-label-results-mysql}
\label{fig:subfig3}
}

\subfloat[MaxDB]{
\includegraphics[width=0.3\textwidth]{figures/multi-label-results-maxdb}
\label{fig:subfig4}
}
\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.
%\todo[inline]{better discussion}% and graphical legend would be nice}
}
\label{fig:mulan}
\end{figure*}

@@

\subsection{Summary of techniques}
While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
supervised learners have the advantage of domain knowledge, which typically means improved results. 
Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.

@@

We observed that MySQL had more topic subsets than MaxDB. 
MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
\todo[inline]{expand}
%\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}

@@

\textbf{Do label frequencies change over time?} -- Yes, MySQL's label
frequencies decreased as that release aged. \emph{Usability} and \emph{reliability}
labels became more and more infrequent as it matured. \emph{Maintainability} topics
became more prevalent as MaxDB matured.

@@

%XXX can just remove this
%\subsection{Effectiveness}

%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, as correctness is a loosely defined concept in common parlance.

@@

\emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages. 
Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.

\emph{Internal validity} -- include inter-rater reliability and our lack of rival explanation building. 
We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
We did not attempt to match our results to any particular model or theory.

@@

% in Case Study Research: Design Case Studies
\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
Unfortunately only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.

@@

Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
They do not use external domain-specific information as we do.
In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.

@@

% para 1 we did this, we were successful
In this paper we presented a cross-project data mining technique, \emph{labelled topic extraction}. The technique leveraged  a non-functional requirements (NFRs) taxonomy to mine and label latent software development topics. %topics and artifacts from project histories.
Our contributions include:
\begin{itemize}
\item We provided an unsupervised method of topic labelling (word-lists);
%\item Our unsupervised method can be applied to projects without training;
\item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
\item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
\item We provided a method of cross-project analysis via topic labelling;
\item We demonstrated a method of visualizing these topics across time.
\end{itemize}

@@

% para 2 we did this to validate and show we did it
% Experiment description -- origin of wordbags
We validated our topic labelling techniques using multiple experiments.
Using unsupervised word-list analysis we set up two NFR topic labelling experiments: 
our \textsf{exp2} experiment used small but accurate word-lists, while \textsf{exp3} used large general word-lists derived from WordNet.
The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
%Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
% we could can improve auto labelling with ML
Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
Both kinds of learners performed well with area under the ROC curve values between $0.6$ and $0.8$.

% para 3 look while we did it we found this
% total topics - maybe move above
We showed that NFRs are often trending topics; that NFRs are quite common in developer topics; and that there are efficient methods of semi-automating and automating topic labelling.
% We found NFRs
Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project. 
% Try to provide an example of NFRs but little context
In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.

@@

%\subsection{Future Work}

 % para 5 future work
 % Future work
There are several avenues of further investigation.  
More external validation would be useful. 
Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail. 
We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, an end-user facing software product.
 % Talk to people
% We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
% It is difficult to map abstract qualities to particular messages. % what?
%  % This is disjointed and needs more context
% On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance? %In this study we take the position that any type of maintenance has to do with software quality.
